HPC--101\\Onboarding

### 2017-02-04 hpc_101/paper.tex
Breakup of sections into tex file

\title[HPC--101]{HPC--101\\Onboarding} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Gwen Jacobs Ph.D.,\\Sean Cleveland Ph.D., Ron Merrill Ph.D.,\\Jennifer Geis M.S., David Schanzenbach M.S.}
\institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Information Technology Services \\
{\ci} \\
University of {\hawaii} \\ % Your institution for the title page
\medskip
\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
\textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address
}
\date{\today} % Date, can be changed to a custom date

@@

\section{Introduction}

\subsection{Parallel Computing}
\begin{frame}
	\frametitle{Parallel Computing}
	\begin{itemize}
		\item High Performance Compute
		\begin{itemize}
			\item Each separate process can send and receive data amongst other processes (MPI \& OpenMP)
			\item If processes must communicate for the overall program to proceed, high-speed networking is needed (only MPI)
		\end{itemize}
		\item High Throughput Compute
		\begin{itemize}
			\item \emph{Pleasantly Parallel} -- Processes are independent and no communication is necessary 
		\end{itemize}
	\end{itemize}
\end{frame}

\section{University of Hawai'i Cluster}

\subsection{{\craycs}}
\begin{frame}
\frametitle{{\craycs} -- History}
\begin{itemize}
 \item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
 \item Delivered in October of 2014
 \item Accepted in December of 2014 
 \item Early adopter testing started in January 2015
 \item Opened for general use in April 2015
 \item March 2016, 92 new nodes added to cluster, raising total core count from 3,800 to 5,876 cores
 \item January 2017, 5 new nodes added to cluster, raising total core count from 5,876 to 5,992 cores
 \item As of Nov 2016, more than 300 users have been granted access to the cluster
\end{itemize}
\end{frame}

\subsubsection{Hardware}
\begin{frame}
	\frametitle{{\craycs} -- Compute Nodes}
	\begin{itemize}
        \item All nodes are diskless with some RAM used for the Operating System
	\item 178 Standard nodes {\tiny(\emph{community})}
	  \begin{itemize}
            {\footnotesize
	    \item Two 10 core {\intel} Ivy-Bridge processors (\emph{20 cores total})
	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
            }
	  \end{itemize}          
	\item 6 Large memory nodes {\tiny(\emph{community})}
	  \begin{itemize}
            {\footnotesize
	    \item Four 10 core {\intel}  Ivy-Bridge processors (\emph{40 cores total})
	    \item 1TB of physical RAM with $\approx1000$GB of useable RAM
            }
	  \end{itemize}			
        \item 1 GPU node {\tiny(\emph{community})}
          \begin{itemize}
            {\footnotesize
	    \item Two 10 core {\intel} Haswell processors (\emph{20 cores total})
	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
            \item 2 Nvidia Tesla K40 GPUs
            }
	  \end{itemize}
        \item 96 Owner nodes
          \begin{itemize}
            {\footnotesize
            \item 33 nodes have two 10 core {\intel} Haswell processors with 256 GB of RAM
            \item 62 nodes have two 12 core {\intel} Haswell processors with 128 GB of RAM
            \item 1 GPU node
            }
          \end{itemize}
	\item CentOS Linux
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage}
	Two storage options are currently available on the {\craycs}
	\begin{enumerate}
		\item {\lustre}
                  \begin{itemize}
                  \item High performance parallel filesystem 
                  \item Available on all the cluster nodes
                    \item Freely available to all users
                  \end{itemize}
		\item ValueStorage
                  \begin{itemize}
                  \item Network attached scale out storage
                  \item Only available from the login nodes
                  \item Available as a for fee service
                  \end{itemize}
	\end{enumerate}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage $\rightarrow$ {\lustre}}
	\begin{itemize}
		\item The {\craycs} has $\approx582$TB of storage space

		\item Primarily used as scratch space for jobs (Input and Output)
		\item User do not have a usage quota (soft or hard)
		\item Certain directories are subject to a 90 day purge policy
		\item \textbf{Data is not backed up!  Users are responsible for their own data}
                \item {\lustre} utilizes RAID 6 arrays, and is fairly robust but \ldots~\\RAID is not a backup
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
        ~\\
	\begin{itemize}
		\item $500$TB of scale out storage
		\item Purchased annually in 0.5TB increments 
	\end{itemize}
	\btVFill

	\begin{center}
	\textbf{ValueStorage Pricing}~\\ 
		\resizebox{0.50\textwidth}{!}{%
		\begin{tabular}{l r }
			\toprule
			\textbf{Product} & \textbf{Annual Cost}~~ \\
	                \midrule
			\midrule
			0.5TB & \$65.00\\
                        0.5TB + Replication  & \$130.00 \\
			\bottomrule
		\end{tabular}
		}
		~\\ \href{http://www.hawaii.edu/its/value-storage-pricing/}{More Information}
		~\\	{\footnotesize \emph{All prices are subject to change}}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Network}
	\begin{itemize}
		\item 40Gb Infiniband inter-connects (QDR)
		\begin{itemize}
			\item High speed inter-connect between~\\compute nodes, {\lustre} storage and Login nodes
                        \item low latency ($\approx1.3\mu$s)
			\item Utilizes the \emph{fat tree network topology}
			\item[] \includegraphics[width=0.20\textwidth]{images/Fat_tree_network} \\[-1ex] {\fontsize{3}{4} \selectfont Source: \url{https://en.wikipedia.org/wiki/Fat_tree} } 		
		\end{itemize}
		\item 10Gb login node internet connectivity
		\begin{itemize}
			\item Speed test from UH to CERN clocked transfer speeds up to 2$+$ Gb/s
		\end{itemize}	
	\end{itemize}
\end{frame}

\subsubsection{Layout}
\begin{frame}
	\frametitle{{\craycs} -- Layout}
	\includegraphics[width=1\textwidth]{images/layout}
\end{frame}

\section{Sustainability}
\subsection{Community Resources}
\begin{frame}
	\frametitle{Community Resources}
	\begin{itemize}
		\item The initial investment in the cluster provides resources for free to all faculty, staff, and students affiliated with the University of {\hawaii}
		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, exclusive.q, lm.q, gpu.q, sb.q, kill.q, htc.q}
		\item For some users, the publicly available resources may not be enough~\ldots
	\end{itemize}
\end{frame}

\subsection{Ownership Models}

\subsubsection{Condo Model}
\begin{frame}
	\frametitle{Condo Model}
	\begin{itemize}
		\item The Condo model allows users to buy nodes (\emph{condos}) to incorporate into the cluster	
		\item Node owners are provided with priority access to their purchased hardware
		\item All nodes have a \emph{5 year warranty}
		\begin{itemize}
			\item Once a nodes warranty has expired, it will be removed from the cluster
		\end{itemize}
		\item Condo owners are given early access to purchased nodes
		  \begin{itemize}
                    \item Access is granted as soon as the funds land in our accounts
		    \item The 5 year warranty will not begin until newly ordered nodes are installed
		\end{itemize}
		\item Condo owners are also given the option to purchase~\\1TB of {\lustre} storage per node purchased
	\end{itemize}
\end{frame}

\subsubsection{Service Units}
\begin{frame}	
	\frametitle{Service Units}
	\begin{itemize}
	 \item In some cases, users will not require owning a node, but still need priority access
	 \item An alternative to purchasing a node, is to purchase service units (\emph{SU})
	 \item SUs come in two flavors:
		 \begin{itemize}
			 \item[--] Standard node units 
				\begin{itemize}
					\item 20 core hours, with access to 128GB of ram on a standard~node
				\end{itemize}
			 \item[--] Large memory node units
				\begin{itemize}
					\item 40 core hours, with access to 1TB of ram on a large~memory~node
				\end{itemize}
		 \end{itemize}
	 \item A minimum order totaling \$500 is required
	\end{itemize}
\end{frame}

\subsubsection{Price Card}
\begin{frame}
	\frametitle{Price Card}
	\begin{center}
	{\large \textbf{Condo Price Card}} \\
	\resizebox{\textwidth}{!}{%
		\begin{tabular}{l r ||l r}
			\toprule
			\textbf{Product} & \textbf{Cost} & \textbf{Product} & \textbf{Cost} \\
			\midrule
			\midrule
			Standard & \$6,600.00  & Standard + 2 GPUs (Nvidia{\regtrademark} K40) & \$13,600.00 \\
			\hline
			Large Memory & \$33,900.00  & 1 TB {\lustre} Storage for 5 years & \$$600.00$ \\
			\bottomrule
		\end{tabular}
		}
	\end{center}
	\smallskip
	\begin{center}
		{\large \textbf{Service Unit Price Card}}\\
		\begin{tabular}{l || c || r}
			\toprule
			\textbf{Product} & \textbf{Cost} & \textbf{Minimum Order}\\
			\midrule
			\midrule
			Standard node & \$0.50 per SU &  1,000~SU (\$500.00)\\
			Large memory node & \$2.00 per SU & 250~SU (\$500.00)\\
			\bottomrule
		\end{tabular}
		\btVFill
	{\footnotesize \emph{All prices are subject to change}}
	\end{center}
\end{frame}

\section{Cluster Interaction}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item{Connecting to a cluster @ UH}
			\begin{itemize}
			\item Login to the cluster
			\item Verify user permissions
			\end{itemize}
    {\semitransp[25]{\item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{Connecting to a cluster @ UH}
\begin{frame}
	\frametitle{Connecting to a cluster @ UH}
	\begin{description}[\setlength{\leftmargini}{0pt}]
	\item[] \begin{itemize}
		\item To connect to the cluster, an SSH client is required
		\item Linux and MacOS X, typically have a SSH client already installed
		\item Windows typically needs a 3rd party SSH client
		\item Suggested SSH clients for Windows include:
		\begin{itemize}
			\item \href{http://www.hawaii.edu/askus/685}{SSH Secure Shell}~(SSH 3.2.9)
			\item \href{http://www.chiark.greenend.org.uk/~sgtatham/putty/download.html}{Putty}
                        \item \href{http://mobaxterm.mobatek.net/}{MobaXterm}
		\end{itemize}
	\end{itemize}
	\item[] ~\\
	\item[] \begin{itemize}
		\item The {\craycs} has two login nodes:
		\begin{itemize}
			\item uhhpc1.its.hawaii.edu
			\item uhhpc2.its.hawaii.edu
		\end{itemize}
	\end{itemize}
	\end{description}
        ~\\~\\
	{\large Let's attempt to login!}
\end{frame}

\subsubsection{Login to the {\craycs}}
\begin{frame}
\frametitle{Login to the {\craycs}}
	\begin{center}\textbf{Windows}\end{center}
	\hrule~\\
	\begin{itemize}
		\item If SSH 3.2.9 installed (Lab PCs have it installed)
		\item Open the start menu, and type ``SSH'' and you should see a program called ``SSH Secure File Terminal Client''
		\item Click ``Quick Connect'' and enter the following information:
			\begin{itemize}
			\item[] \textbf{Host Name:} uhhpc1.its.hawaii.edu --OR-- uhhpc2.its.hawaii.edu
			\item[] \textbf{User Name:} Your UH User name e.g., user99
			\item[] \textbf{Port:} 22
			\end{itemize}
		\item Press ``Connect''
		\item Enter your UH user password when prompted and press the return key
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Login to the {\craycs}}
	\begin{center}\textbf{Mac \& Linux}\end{center}
	\hrule~\\
	\begin{itemize}
		\item Open a terminal window
		\item Enter one of the following:
		\begin{itemize}
			\item ssh $<$UH User name$>$@uhhpc1.its.hawaii.edu
			\item ssh $<$UH User name$>$@uhhpc2.its.hawaii.edu
			\item \textbf{Example:} ssh user99@uhhpc1.its.hawaii.edu
		\end{itemize}
		\item Enter your UH user password when prompted and press the return key
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{On Initial Login~\ldots}
	Validate that all system permissions are correct for your user~\\
	\begin{enumerate}
		\item Test that you can list files in your home: `\textcolor{blue}{\texttt{ls -la}}'    
		\item Test making a file in your home: `\textcolor{blue}{\texttt{touch test.txt}}'    
		\item Go into \ctilde{}/lus: `\textcolor{blue}{\texttt{cd \ctilde{}/lus/}}'  
		\item Test making a file in your lus directory: `\textcolor{blue}{\texttt{touch test.txt}}'   
		\item Go into \ctilde{}/apps:  `\textcolor{blue}{\texttt{cd \ctilde{}/apps/}}'     
		\item Test making a file in your apps directory: `\textcolor{blue}{\texttt{touch test.txt}}'   
	\end{enumerate}
	\begin{alertblock}{Result}
		\begin{center}Did you get any errors?  Let us know if you did\end{center}
	\end{alertblock}
	\btVFill

	\small Notes:
		\begin{itemize}\tiny
		\item On login you are placed in /home/$<$username$>$/
		\item By default, \ctilde{} is equivalent to /home/$<$username$>$/
		\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
    \item User directories
		{\semitransp[25]{\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{User Directories}
\begin{frame}[fragile]
\frametitle{User Directories}
\begin{block}{Home}
\begin{semiverbatim}\tiny \texttt
[user99@login \ctilde]\$ ls -l 
total 0
lrwxrwxrwx 1 user99 user99 23 Jan 15 20:38 \textcolor{teal}{apps} -> /lus/scratch/usr/user99
lrwxrwxrwx 1 user99 user99 19 Jan 15 20:38 \textcolor{teal}{lus} -> /lus/scratch/user99
lrwxrwxrwx 1 root   root   37 Jan 15 20:41 \colorbox{black}{\textcolor{red}{purge}} -> /lus/scratch/log/purge/current/user99
\end{semiverbatim}
\end{block}
	\begin{itemize}\footnotesize
		\item \ctilde{}/ is not on the {\lustre} filesystem and \textbf{should not be used for job data!}
		\item \ctilde{}/lus/ is a symlink to the {\lustre} scratch
		\begin{itemize}\tiny
			\item This is where all your job data files should live
			\item Items in this directory \emph{\textbf{are}} subject to our 90 day purge policy
		\end{itemize}
		\item \ctilde{}/apps/ is a symlink to where programs should be stored
		\begin{itemize}\tiny
			\item Items in this directory \emph{\textbf{are not}} subject to our 90 day purge policy
			\item Directory is monitored for abuse
		\end{itemize}
		\item \ctilde{}/purge/ is typically a dead symlink
		\begin{itemize}\tiny
			\item Symlink becomes active if the user has files that are part of the next automatic purge
			\item When \ctilde{}/purge/ is active, the directory containing two files -- \emph{purge\_list.txt} \& \emph{totals.txt}
			\item An email is sent to users if they have files that will be purged
			\item Email notification is sent out 14 days before the purge takes place
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{User Directories}
\begin{block}{Filesystems}
\begin{semiverbatim}\tiny \texttt
[user99@login \ctilde]\$ df -h
Filesystem                                  Size  Used Avail Use% Mounted on
10.10.0.3:/ha_cluster/home                  1.8T  888G  851G  52% /home
10.12.0.51@o2ib:10.12.0.52@o2ib:/scratch    582T  429T  125T  78% /lus/scratch
\end{semiverbatim}
\end{block}
	\begin{itemize}\footnotesize
		\item /home/$<$username$>$ exists on a NFS mounted filesystem
		\begin{itemize}\tiny
			\item Only has 1.8TB of useable space
			\item Using all this space may cause problems for the entire cluster
			\item Not a high performance filesystem and small in size
		\end{itemize}
		\item /lus/scratch/ is the {\lustre} filesystem
		\begin{itemize}\tiny
			\item Has 582TB of useable space
			\item \ctilde{}/apps/, \ctilde{}/lus/, and \ctilde{}/purge/ all point to directories on this filesystem
			\item High performance and a lot more space for users to use
			\item No hard or soft quotas are in place
			\item Utilization is managed through the 90 day purge policy
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
    {\semitransp[25]{\item User directories}}
		\item Transferring files
		\begin{itemize}
			\item Globus
		\end{itemize}
		{\semitransp[25]{
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions			
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{Transferring Files}
\begin{frame}
	\frametitle{Available File Transfer Protocols}
	\begin{itemize}
		\item The cluster has the following options for transferring files:
		\begin{itemize}
			\item[--] scp (RCP$+$SSH protocol)
			\item[--] rsync (rsync protocol with SSH transport)
			\item[--] SFTP (SSH FTP protocol) -- Filezilla, Cyberduck
			\item[--] \href{https://www.globus.org/}{Globus} (Grid FTP protocol)
		\end{itemize}
		\item All options are widely used, and have clients that can be found for on most major operating systems
%               \item Please see the the CI website for links to several different clients?                 
	\end{itemize}
	\btVFill
	\begin{center}
		SFTP, scp, and rsync are fairly common on Linux systems,~\\but Globus is not as common~\ldots
	\end{center}

\end{frame}

\subsection{Globus}
\begin{frame}
	\frametitle{Globus}
	\begin{block}{What is Globus?}\footnotesize
	The Globus transfer service provides high-performance, secure, file transfer and synchronization between endpoints.~\\~\\
	Globus handles all the difficult aspects of data transfer, allowing application users to easily start and manage transfers between endpoints, while automatically tuning parameters to maximize bandwidth usage, managing security configurations, providing automatic fault recovery, and notifying users of completion and problems. 
	\end{block}
	\begin{definition}\tiny
	An \textbf{\emph{endpoint}} is one of the two file transfer locations -- either the source or the destination -- between which files can move.~\\Once a resource (such as a server, cluster, storage system, laptop, or other system) is defined as an endpoint, it will be available to authorized users who can transfer files to or from this endpoint.
	\end{definition}
\numlessfootnotetxt{\tiny \url{https://www.globus.org/file-transfer}}
\end{frame}

\begin{frame}
	\frametitle{Globus}\footnotesize
\begin{block}{How do I get Globus?}
To utilize Globus, follow the following steps:
\begin{itemize}\tiny
\item Register for a Globus Online account -- \url{https://www.globusonline.org/signup}
\item Sign in to Globus Online (using your Globus Online username and password) -- \url{https://www.globusonline.org/signin}
\item Select ‘Start Transfer’ under ‘File Transfer’, or from the drop down menu in the top bar
\item You can view the list of available endpoints by clicking the button on the ‘Endpoint’ drop down box
\begin{itemize}\tiny
 \item Each of the login nodes is also an endpoint: \textbf{\emph{UH-HPC1}} \& \textbf{\emph{UH-HPC2}}
\end{itemize}
\item Once you select an endpoint, a login window will pop up. You can access the UHHPC endpoints by simply using your UH username and password. Enter your UH accounts username in the ‘Username’ field and UH accounts password in the ‘Password’ field and click ‘Authenticate’. 
\item You will see a listing of the contents of your home directory on the UH HPC. Double click on a directory to view its contents
\item Select a file or directory and click on the highlighted ‘arrow button’ to initiate the transfer
\end{itemize}
\end{block}
~\\
These instructions can also be found on the {\ci} website:~\\
{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
\end{frame}

\begin{frame}
	\frametitle{Globus}
        \begin{itemize}
          \item In order to transfer data from the {\craycs} to your personal computer, a client called the Globus Connect Personal (\url{https://www.globus.org/globus-connect-personal}) needs to be installed
          \item The Globus Connect Personal, turns your personal computer into a private endpoint that is only useable with your personal Globus Online account.
          \item If you want to install Globus on your own server, Globus Connect (\url{https://www.globus.org/globus-connect-server}) is required
            \item If issues arise while trying to install Globus Connect or Globus Connect Personal,  please contact us and we will be more than happy to help
\end{itemize} 
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
   {\semitransp[25]{ \item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		}}
		\item Software
		\begin{itemize}
			\item Modules
			\item Acquiring software
			\item Compilers
		\end{itemize}
		{\semitransp[25]{
		\item Managing user jobs
		\begin{itemize}{\semitransp[25]{
			\item Job scheduler
			\item Using SLURM
			\item Partitions
			\item Submitting jobs (Examples)
			}}
		\end{itemize}			
		}}
	\end{itemize}
\end{frame}

\subsection{Software}
\begin{frame}
	\frametitle{Modules}
	\begin{block}{Modules}\tiny
	A tool to help users manage their Unix or Linux shell environment, by allowing groups of related environment-variable settings to be made or removed dynamically.\footnotemark

	\end{block}
	\begin{itemize}
		\item We globally install frequently requested software packages and create modules for all users to access
		\item Access to modules is via the \textbf{module} command
		\begin{itemize}\footnotesize
			\item `module avail' -- list installed modules
			\item `module show $<$module name$>$' -- Show what actions a module performs
			\item `module load $<$module name$>$' -- Loads the named module
			\item `module purge' -- Unload all loaded modules
		\end{itemize}
		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
	\end{itemize}
	\footnotetext[1]{\tiny \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
\end{frame}

\begin{frame}
	\frametitle{Acquiring Software -- Binaries and/or Source}
	\begin{itemize}
		\item	You can transfer software source, binaries or scripts into your \ctilde{}/apps directory on the {\craycs}
		\begin{itemize}
			\item Binaries compiled as x86\_64 (64-bit) for CentOS 6.5 or RHEL6.5 should work
		\end{itemize}
		\item You may also download tar or zipped software/source code directly from the login nodes using tools like \textbf{wget} \& \textbf{curl}
		\item You may also clone source repositories using the correct software revision tool: \textbf{git}, \textbf{svn}, \textbf{hg}, \textbf{cvs}, etc.
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Compilers}
	\begin{itemize}
		\item We have the {\intel}, GNU (gcc, g$++$), {\cray} \& PGI{\regtrademark} compilers
		\item Compiling must take place on a compute node
		\begin{itemize}
			\item Interactive sessions are useful for compiling software
			\item Sandbox nodes mirror the environment the compute nodes provide and are ideal for compilation
			\item Login nodes \textbf{do not} load all the software and libraries found on the compute nodes
		\end{itemize}
		\item {\intel} compilers are recommended for best performance
		\begin{itemize}
			\item {\intel} 2013 compilers:
			\begin{itemize}
				\item module load intel/ics -- Loads {\intel} compilers: {\tiny \textbf{icc}, \textbf{ifort}, \textbf{icpc}}
				\item module load intel/impi -- Loads {\intel} MPI wrapper: {\tiny\textbf{mpiicc}, \textbf{mpiifort}, \textbf{mpiicpc}}
			\end{itemize}
			\item {\intel} 2016 compilers:
			\begin{itemize}
				\item We have 2 floating seats for {\intel} 2016 compiler
				\item intel\_2016/ics 
				\item intel\_2016/impi 
			\end{itemize}
		\end{itemize}			
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Overview -- Cluster Interaction}
	\begin{itemize}
		\item {\semitransp[25]{Connecting to a cluster @ UH}}
			\begin{itemize}{\semitransp[25]{
			\item Login to the cluster
			\item Verify user permissions
			}}
			\end{itemize}
   {\semitransp[25]{ \item User directories
		\item Transferring files
		\begin{itemize}{\semitransp[25]{
			\item Globus
		}}\end{itemize}
		\item Software
		\begin{itemize}{\semitransp[25]{
			\item Modules
			\item Acquiring software
			\item Compilers
			}}
		\end{itemize}
		}}
		\item Managing user jobs
		\begin{itemize}
			\item Job scheduler
			\item Using SLURM
			\item Partitions
			\item Submitting jobs (Examples)
		\end{itemize}		
	\end{itemize}
\end{frame}

\subsection{Managing User Jobs}
\begin{frame}
\frametitle{Managing User Jobs}
User jobs all come in different shapes and sizes:  
\begin{itemize}\footnotesize
	\item Require multiple nodes working in concert towards a common goal (MPI)
	\item Require a single node, in which they use multiple threads~\\work together (OpenMP, pthreads)
	\item Require a lot of cores to process a lot of data in an identical manner,~\\yet none of the inputs have dependencies on another (HTC)
\end{itemize}
\bigskip
The {\craycs} is capable of handling many different types of jobs, but with so many users in a multi-user environment, how do we impose order on this chaos?
\btVFill
\begin{center}This looks like a job for a \textbf{\emph{job scheduler}}!\end{center}
\end{frame}

\begin{frame}
  \frametitle{Job Schedulers}
  \begin{block}{Purpose}To control and prioritize the execution order of unrelated jobs\end{block}
	Basic features expected of a job scheduler:
	\begin{itemize}\footnotesize
        \item Provides a user interface for users to request resources and monitor work 
        \item Allocates access to resources for the requested duration of time
	\item Starts, monitors and terminates work on allocated resources
	\item Arbitrates contention for resources by managing queues of pending work
	\end{itemize}
	\bigskip
        The {\craycs} uses the \textbf{S}imple \textbf{L}inux \textbf{U}tility for \textbf{R}esource \textbf{M}anagement~\\ or simply known as the \emph{SLURM scheduler}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
	\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/slurm.html}}
\end{frame}

\subsection{SLURM}
\begin{frame}
\frametitle{SLURM}
\begin{block}{How are jobs scheduled?}
User submitted jobs are assigned a priority using a fairshare algorithm.\\
Factors such as the following are all used to assign a priority to a given job: 
\begin{itemize}
\item Runtime
\item Amount of resources requested
\item Age of job
\item Amount of core hours a user has used in recent history
\end{itemize}

\end{block}
\end{frame}

\begin{frame}
\frametitle{SLURM commands}\footnotesize
SLURM has a series of commands, each of which allow users to interact with the job scheduler
\begin{itemize}\tiny
	\item \emph{\textbf{srun}} --  Used to submit a job for execution or initiate job steps in real time
	\item \emph{\textbf{sbatch}} -- Used to submit a job script for later execution. The script could contain one or more srun commands
	\item \emph{\textbf{squeue}} -- Reports the state of jobs or job steps
	\item \emph{\textbf{scancel}} -- Used to cancel a pending or running job or job step. It can also be used to send an arbitrary signal to all processes associated with a running job or job step
	\item \emph{\textbf{sinfo}} -- Reports the state of partitions and nodes managed by Slurm. It has a wide variety of filtering, sorting, and formatting options
	\item \emph{\textbf{sacct}} -- Used to report job or job step accounting information about active or completed jobs
	\item \emph{\textbf{scontrol}} -- The administrative tool used to view and/or modify Slurm state. Note that many scontrol commands can only be executed as user root
\end{itemize}
\begin{itemize}
	\item[--] Examples usage of the SLURM commands can be seen on schedmd's \href{http://slurm.schedmd.com/quickstart.html}{quickstart}
	\item[--] Each command should have a `man' page, or displays help when the -h flag is used
\end{itemize}
\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
\end{frame}

\subsection{SLURM Partitions}
\begin{frame}
\footnotesize
\frametitle{Partitions}
\begin{block}{What is a partition?}\footnotesize
A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.  Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc. Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
\end{block}
\begin{itemize}
	\item The {\craycs} currently has seven public partitions:~\\\textbf{community.q}, \textbf{exclusive.q}, \textbf{lm.q}, \textbf{gpu.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
\end{itemize}
\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
\end{frame}

\begin{frame}
\frametitle{Partitions}
\includegraphics[width=0.95\textwidth]{images/partitions}
\end{frame}

>>>>

\include{sections/00_title}

@@

\include{sections/01_intro}
\include{sections/02_cluster}
%\include{sections/03_sustainability}
\include{sections/04_cluster_interaction}
\include{sections/05_reporting}
\include{sections/06_ediquette}
\include{sections/07_policies}
\include{sections/08_end}
%\include{sections/09_related_services}
\include{sections/10_faq}
